---
title: "Post01-Matt-Brennan"
author: "Matt Brennan"
date: "10/29/2017"
output: github_document
---
Hello fellow bloggers,

Introduction:

Today I am going to talk about one of the most useful packages in the R library, sqldf! Although the R language enables the programmer to perform various mathematical functions with data, there oftentimes is a disconnect between this language and various others. The primary reason for this divide is that most data analysts in businesses deal with relational database systems (RDBMS), where the R language fails to incorporate these systems.

Motivation:

The fix: the ambitious package sqldf!

Through this blog, I hope to not only demonstrate the effectiveness of this package, but captivate the reader into utilizing sqldf in future endeavors. The sqldf package allows one to use R but through SQL commands. Sprinkled throughout the blog will be examples, interesting facts, and descriptions of the actions that are being performed!

In order to install the package one must use the code below.
```{r}
install.packages("sqldf")
```
After the package is installed, one must library or load the package.
```{r}
library(sqldf)
```

For the examples in this blog we will incoporate the stats of all nba players during the 2017 season. One can find this data in the data folder within post01.
```{r}
github <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/"
csv <- "data/nba2017-stats.csv"
download.file(url = paste0(github, csv), destfile = 'nba2017-stats.csv')
dat <- read.csv('nba2017-stats.csv', stringsAsFactors = FALSE)
```
```{r}
dat
```

Examples:

Ex 1: Select the player column only from dat!
```{r}
sqldf("SELECT 
      player 
      FROM dat")
```
Ex 2: Select the columns, player, and games_played from dat but change the name of games_played to gp!
```{r}
sqldf("SELECT
      player
      , games_played as gp
      FROM dat")
```
Lets add a column using familiar dplyr to create a points column for further study.
```{r}
library(dplyr)
```
```{r}
dat <- dat %>%
      mutate(points = 3*dat$points3_made + 2*dat$points2_made + dat$points1_made)
```
Ex 3: Construct a new data frame with only the player and the points columns, because points is the only stat that matters really matters.
```{r}
dat_points_only <- sqldf("SELECT
                         player
                         , points
                         FROM dat")
dat_points_only
```
As referenced by R-bloggers, "R users reshape data frames while database users construct queries and return result sets.  In either case, we are dealing with the same basic idea – essentially a rectangular grid containing a set of values obtained from a data source."

Let's now make a new data table that takes only some of the data provided for simplicity, let's call this dat_stats:
```{r}
dat_stats <- sqldf("SELECT
                   player
                   , points
                   , assists
                   , minutes
                   , games_played
                   FROM dat")
dat_stats
```
Yet, we want more essential data to come from another source also!
```{r}
github <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/"
csv <- "data/nba2017-roster.csv"
download.file(url = paste0(github, csv), destfile = 'nba2017-roster.csv')
dat1 <- read.csv('nba2017-roster.csv', stringsAsFactors = FALSE)
```
```{r}
dat1
```
Ex 4: Unique notation where one can select various columns from the multiple data_frames based on the a similarity in the the columns. For this example we are compiling important information around the name of the player.
```{r}
dat_stats2 <- sqldf("SELECT a.player, a.points, a.assists, b.position, b.age, b.experience, b.salary, b.team
                    FROM dat_stats as a, dat1 as b
                    WHERE a.player = b.player
                   ")
dat_stats2
```
For further information about how to utilize this type of merging, click on the link below to a useful Youtube video!
https://www.youtube.com/watch?v=9JRsHxKCvsg

Ex 5: Using this new data_stats dataframe, lets focus in on only the players that are less than 25
```{r}
sqldf("SELECT *
      FROM dat_stats2
      WHERE age < 25")
```
The Star symbol is a very useful notation for this syntax as it pertains to selecting all of the data: https://www.r-bloggers.com/manipulating-data-frames-using-sqldf-a-brief-overview/ can be used to learn more about this notation.

For individuals that are well adjusted to SQL, yet need to improve upon their work in R, sqldf will be the easiest transitioning mechanism.
Fun fact about SQL:SQL was invented in 1970 by Donald Chamberlin and is the largest database language.

In order to plot certain charts we know would like to utilize ggplot2:
```{r}
library(ggplot2)
```

Ex 6: Let's make a bar-chart of the salary versus points for individuals that are 30 or older in age!
```{r}
setwd("~/Stat133/Stat133-hws-fall17/post01/images")
```

```{r}
dat_pts_sal <- sqldf("SELECT points, salary
                      FROM dat_stats2
                      WHERE age >= 30")
gg_pts_sal <- ggplot(data = dat_pts_sal, aes(x = points, y = salary)) + geom_point() +ggtitle("Points versus Salary")

```
Let's now export this scatterplot to the images folder!
```{r}
getwd()
```
```{r}
ggsave(filename = "points_salary_30s.pdf", plot = gg_pts_sal, width = 6, height = 4)
```


Let's now try to master the Group By aspect of sqldf:

Suppose we want to get all of the individuals making more than 10 mil, and we want to get the count of how many there are on each team. The Group By aspect can be a very useful and easy way to perform this.
```{r}
sal_grouped <- sqldf("SELECT team, COUNT (player) as occurences
                     FROM dat_stats2
                     WHERE salary > 10000000
                     GROUP BY team")
sal_grouped
```
For more information about the GROUP BY command visit http://analyticsplaybook.org/api/learn_sql_30_minutes.html to watch a video.

Let's now upload a histogram of the amounts to images as a pdf in order to keep track of the results:
```{r}
setwd("~/Stat133/Stat133-hws-fall17/post01/images")
```

```{r}
pdf(file = "histogram-salary.pdf", width = 7, height = 5)
hist(sal_grouped$occurences, xlab = "Making over 10 mil.", main = "Number of players making 10 mil. per team")
dev.off()
```
Let's now work with the intersection and EXCEPT aspects of sqldf in order to solidify all of the possible methods by which one could create a dataframe using this package:

Since we have created a variety of data_frames, such as dat_stats2 which we constructed from merging dat_stats and dat1, we can now work backwards to see what these original groups have in common with the newer dat_stats2:

Let's create a new data_frame, dat_sal_gt_10mil, consisting of only players that make more than 7 mil!
```{r}
dat_sal_gt_10mil <- sqldf("SELECT *
                          FROM dat_stats2
                          WHERE salary > 10000000")
dat_sal_gt_10mil
```
Now, let's use Intersect in order to see what it does:

```{r}
sqldf("SELECT player, points
      FROM dat_sal_gt_10mil
      INTERSECT
      SELECT player, points
      FROM dat_stats")
```
You see that the results are only the player, points from dat_stats that are in common between the two data_frames. In the next example one can see that the switching up the order should give the exact same results:
```{r}
sqldf("SELECT player, points
      FROM dat_stats 
      INTERSECT
      SELECT player, points
      FROM dat_sal_gt_10mil")
```

Furthermore, one can utilize the EXCEPT command to identify the rows that the first data_frame has that the second data_frame does not:

```{r}
sqldf("SELECT player, points
      FROM dat_sal_gt_10mil
      EXCEPT
      SELECT player, points
      FROM dat_stats")
```
It is not surprising that the results of the previous command did not provide any rows as the first data_frame listed was a sub-frame of the second one. The next example should provide more relevant information as this command comprises all of the rows that are in the larger data_frame that do not appear in the smaller one.
```{r}
sqldf("SELECT player, points
      FROM dat_stats
      EXCEPT
      SELECT player, points
      FROM dat_sal_gt_10mil")
```
The intersection and except commands can be very integral parts of any data manipulation. With this being said, for more information about how these elements work utilize https://nishanthu.github.io/articles/SQLSetandJoin.html to learn more.


Take Home Message:
Anyone that has utilized the SQL language in previous courses or through work experience, the sqldf function allows one to utilize these already gained skills but via the R language. This would allow one to best understand the differences between the languages and thus allow for the maximum results.



References:
https://www.r-bloggers.com/make-r-speak-sql-with-sqldf/
https://www.youtube.com/watch?v=9JRsHxKCvsg
http://blog.yhat.com/posts/10-R-packages-I-wish-I-knew-about-earlier.html - where I found out about this syntax
https://www.r-bloggers.com/manipulating-data-frames-using-sqldf-a-brief-overview/
http://analyticsplaybook.org/api/learn_sql_30_minutes.html
https://nishanthu.github.io/articles/SQLSetandJoin.html
https://cran.r-project.org/web/packages/sqldf/index.html

